Reliable error estimation for Sobol' indices

### 2016-08-08 article.tex
Swapping section 3.1 with section 3.2
changes L, sections 2.3, 4.1, 4.2 and 3.1 (b^m-1)

\newcommand{\fudge}{\fC}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
%\renewcommand{\bbK}{\natzero^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vx}{\bvec{x}}
\newcommand{\vI}{\bvec{I}}
\newcommand{\vk}{\bvec{k}}
\newcommand{\vz}{\bvec{z}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\bvec{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{iid}{\sim}}

@@

\subsection{Definition of Sobol' indices}
We adopt the same notations introduced by Owen in \cite{Owen}. Denote by $f$ a numerical model, by $\vx=(x_1,\dots,x_d)$ its vector of inputs and set $\mathcal{D}=\{1,\dots,d\}$. The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$. Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vz=(\vx_u:{\vx'}_{-u})$ is defined as $z_j=x_j$ if $j \in u$ and $z_j={x'}_j$ if $j \notin u$.

@@

\subsection{Estimation of Sobol' indices}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. 
\bigskip

@@

\subsection{Towards a robust estimation}
The aim of this paper is to propose a recursive estimation procedure to estimate Sobol' indices. The choice of the stopping criterion is the key point of this paper. The two estimation procedures reviewed in the previous section can be rendered recursive. In \cite{Gilquin16}, the authors propose a recursive version of the replication procedure. The stopping criterion of the recursive algorithm is an absolute difference between estimates computed on consecutive steps.

More generally, in most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly upon the estimates. Such stopping criteria often involved hyperparameters hard to tweak but above all fail to guarantee any error bound on the estimation. 
\bigskip

Our recursive procedure stands apart from others approaches by the construction of a robust stopping criterion. This criterion is an error bound based on the discrete Walsh decomposition of integrals (\ref{first.order}) and (\ref{total.effect}). This decomposition use the close set property of digital sequences. As such, our recursive procedure relies on the iterative construction of Sobol' sequence.

Background on Sobol' sequences are provided in \cite{crass}. The iterative construction of the Sobol' sequences is performed accordingly to the multiplicative  approach presented in \cite{crass}. The construction of our stopping criterion is the subject of the next section \ref{section.error}. Our recursive estimation procedure is presented in section \ref{section.recursive.procedure}.

\section{Reliable Error Estimation for Cubatures}
\label{section.error} 
\subsection{Definition of $\widehat{S}$ (fix it with max and min)}
In some applications, it is more useful to estimate the normalized Sobol' indices, $S_u = \underline{\tau}_u^2/\sigma^2$ and $\overline{S}_u = \overline{\tau}_u^2/\sigma^2$. In both cases, regular and normalized indices can be approximated by estimating some integrals,
\begin{align*}
S_u & = \frac{\int_{[0,1)^{2d-1}} f(\vx)f(\vx_{u}:{\vx'}_{-u})d\vx d{\vx'} - \mu^2 }{\int_{[0,1)^{d}} f(\vx)^2 d{\vx}-\mu^2} = \frac{I_1-(I_3)^2}{I_2-(I_3)^2}, \\
\overline{S}_u & = \frac{\frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2d\vx d{\vx'}}{\int_{[0,1)^{d}} f(\vx)^2 d{\vx}-\mu^2} = \frac{I_1}{I_2-(I_3)^2}.
\end{align*}
Therefore, we can define $S_u(\vI)$ and $\overline{S}_u(\vI)$ as functions over a vector of integrals $\vI$. If we can estimate $\vI$ such that $\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI}):=[\widehat{\vI}-\varepsilon_{\vI},\widehat{\vI}+\varepsilon_{\vI} ]$, then we can define the following estimators for the Sobol' indices:
\begin{align*}
\widehat{S}_u & = \frac{1}{2}\left(\min\left(\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),1\right) + \max\left(\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),0\right) \right) \\
\varepsilon_{S_u} & = \frac{1}{2}\left(\min\left(\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),1\right) - \max\left(\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),0\right) \right)
\end{align*}
where the true value $S_u\in [\widehat{S}_u - \varepsilon_{S_u}, \widehat{S}_u + \varepsilon_{S_u}]$.

@@

We assume we have an embedded sequence of digital nets in base $b$,
\[
\mathcal{P}_0=\{\bvec{0}\}\subset\dots\subset\mathcal{P}_m=\{\vx_i\}_{i=0}^{b^m}\subset\dots\subset\mathcal{P}_\infty=\{\vx_i\}_{i=0}^{\infty}.
\]

@@

\section{Recursive estimation procedure}
\label{section.recursive.procedure}
%
%$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each procedure.
%\begin{algorithm}[!ht]
%\caption{Procedure $A$}
%\begin{algorithmic}[1]
%\vspace*{0.2cm}
%\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
%\While {stopping criterion}
%\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$
%
%\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
%\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
%\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=2^{\ell-1}}^{2^{\ell}-1}$
%\State Compute  $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec})
%\EndFor
%\State $\ell \leftarrow \ell + 1$
%\EndWhile
%\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
%\end{algorithmic}
%\label{procedureA}
%\end{algorithm}
%\begin{algorithm}[!ht]
%\caption{Procedure $B$}
%\begin{algorithmic}[1]
%\vspace*{0.2cm}
%\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
%\While {stopping criterion}
%\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$
%
%\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
%\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\For {$u \in \mathcal{D}$}
%\State  order $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec})
%\EndFor
%\State $\ell \leftarrow \ell + 1$
%\EndWhile
%\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
%\end{algorithmic}
%\label{procedureB}
%\end{algorithm}
%
%\subsection{Improvements}
%This section discusses two improvements to ease the estimation of Sobol' indices with procedures $A$ and $B$. 
%
%\bigskip
%
%The first improvement focus on the use of a better estimator to evaluate small first-order Sobol' indices with procedure $A$. This estimator is called ``Correlation 2" and was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices. Our aim is to show that using ``Correlation 2" in procedure $A$ may reduces the total number of model evaluations. ``Correlation 2" formula writes as follows:
%\begin{equation}
%\widehat{\underline{\tau}_u^2} = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i))
%\label{correlation2}
%\end{equation}
%It uses an extra set of $n$ model evaluations to estimate $\widehat{\underline{\tau}_u^2}$. Owen explains the effectiveness of ``Correlation 2" as follows. If the main effect of $\vx_u$ is small then $f(\vx)$ will be almost entirely explained by $\vx_{-u}$. The two terms $(f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))$ and $(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i))$ should then be small quantities as well as (\ref{correlation2}). 
%
%We discuss now the potential improvement brought by the use of ``Correlation 2". Assume that the number of small first-order indices is known and equals $\alpha$. Then, the cost of procedure $A$ including ``Correlation 2" to estimate small indices writes: 
%\[ n(d-\alpha) + 2 n'\alpha + 2\max(n,n'),\]
%where: \begin{itemize}
%\item[$\bullet$] $n(d-\alpha)$ is the number of evaluations induced by the $d-\alpha$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i\geq0}$ to estimate all non small indices, 
%\item[$\bullet$] $2 n'\alpha$ is the number of evaluations induced by the $2\alpha$ designs $\{\vz_{i,u}:\vx_{i,-u}\}_{i\geq0}$ and $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i\geq0}$ to estimate all small indices,
%\item[$\bullet$] $2\max(n,n')$ is the number of evaluations  induced by the $2$ designs $\{\vx_{i}\}_{i\geq0}$ and $\{{\vx'}_{i}\}_{i\geq0}$.
%\end{itemize}
%At the opposite, the cost of procedure $A$ excluding ``Correlation 2" writes: 
%\[ n(d-\alpha) + n\alpha + 2n.\]
%It is straightforward that procedure $A$ including ``Correlation 2" performs equally well or better when $n' \leq 2n$. Furthermore, since we are working with Sobol' sequences $n'=2^{k'}$ and $n=2^{k}$. Hence, procedure $A$ including ``Correlation 2" performs equally well or better when the recursive estimation procedure finished at least one step sooner. Illustrations of this improvement will be provided in Section \ref{appli}.
%
%\bigskip
%
%The second improvement focus on the choice of the $2d$ generating matrices that generate the two Sobol' sequences $\mathcal{P}_\ell$  and ${\mathcal{P}'}_\ell$ in procedure $B$.
%
%We note by $C_1,\dots,C_d$ the matrices generating $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and ${C'}_1,\dots,{C'}_d$ those generating $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$.
%In the multiplicative approach \cite{crass}, these $2d$ matrices are selected as the first $2d$ elements of the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values. 
%
%%In procedure $A$, each design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i\geq0}$ constructed in step $6$ of Algorithm \ref{procedureA} is a $d$-dimensional Sobol' sequence generated by the $d$-subset: $${C'}_1,\dots,{C'}_{u-1},C_u,{C'}_{u+1},\dots,C_d.$$ Therefore, $2$-dimensional projection $t$-values are still optimized.
%
%In procedure $B$, each permutation $\pi_u$ applied in step $6$ of Algorithm \ref{procedureB} results in a design $\{\vx_{i,u}:{\vx'}_{\pi_u(i),-u}\}_{i=0}^{2^\ell-1}$. The $d$ generating matrices corresponding to this design are:
%$${C'}_1U_u,\dots,{C'}_{u-1}U_u,C_u,{C'}_{u+1}U_u,{C'}_dU_u$$ where $U_u$ is a a full rank upper triangular matrix verifying $C_u={C'}_uU_u$. The right multiplication by $U_u$ is the digital representation of the ordering by $\pi_u$.

@@

>>>>

\newcommand{\fudge}{\fC}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
%\renewcommand{\bbK}{\natzero^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vx}{\bvec{x}}
\newcommand{\vI}{\bvec{I}}
\newcommand{\vk}{\bvec{k}}
\newcommand{\vw}{\bvec{w}}
\newcommand{\vz}{\bvec{z}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\bvec{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{iid}{\sim}}

@@

\subsection{Definition of Sobol' indices}
%We adopt the same notations introduced by Owen in \cite{Owen}. 
Denote by $f$ a numerical model, by $\vx=(x_1,\dots,x_d)$ its vector of inputs and set $\mathcal{D}=\{1,\dots,d\}$. The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$. Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vw=(\vx_u:{\vx'}_{-u})$ is defined as $w_j=x_j$ if $j \in u$ and $w_j={x'}_j$ if $j \notin u$.

@@

\subsection{Estimation of Sobol' indices}
\label{estimation.strategies}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. 
\bigskip

@@

\subsection{Towards a reliable estimation}
The aim of this paper is to propose a recursive estimation procedure to estimate Sobol' indices. The choice of the stopping criterion is the key point of this paper.

The two estimation procedures reviewed in the previous section can be rendered recursive. In \cite{Gilquin16}, the authors propose a recursive version of the replication procedure. In this recursive version, the stopping criterion is an absolute difference between estimates computed on consecutive steps. More generally, in most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly upon the estimates. Such stopping criteria often involved hyperparameters hard to tweak but above all fail to guarantee any error bound on the estimation. 
\bigskip

Our recursive procedure stands apart from others approaches by the construction of a robust stopping criterion. This criterion is an error bound based on the discrete Walsh decomposition of integrals (\ref{first.order}) and (\ref{total.effect}). This decomposition use the close set property of digital sequences. As such, our recursive procedure relies on the iterative construction of Sobol' sequences. The iterative construction of the Sobol' sequences is performed accordingly to the multiplicative  approach presented in \cite{crass}.

The construction of our stopping criterion is the subject of the next section \ref{section.error}. Our recursive estimation procedure is presented in section \ref{section.recursive.procedure}.

\section{Reliable Error Estimation for Cubatures}
\label{section.error}

@@

We assume we have an embedded sequence of digital nets in base $b$,
\[
\mathcal{P}_0=\{\bvec{0}\}\subset\dots\subset\mathcal{P}_m=\{\vx_i\}_{i=0}^{b^m-1}\subset\dots\subset\mathcal{P}_\infty=\{\vx_i\}_{i=0}^{\infty}.
\]

@@

\subsection{Definition of $\widehat{S}$ (fix it with max and min)}
In some applications, it is more useful to estimate the normalized Sobol' indices, $S_u = \underline{\tau}_u^2/\sigma^2$ and $\overline{S}_u = \overline{\tau}_u^2/\sigma^2$. In both cases, regular and normalized indices can be approximated by estimating some integrals,
\begin{align*}
S_u & = \frac{\int_{[0,1)^{2d-1}} f(\vx)f(\vx_{u}:{\vx'}_{-u})d\vx d{\vx'} - \mu^2 }{\int_{[0,1)^{d}} f(\vx)^2 d{\vx}-\mu^2} = \frac{I_1-(I_3)^2}{I_2-(I_3)^2}, \\
\overline{S}_u & = \frac{\frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2d\vx d{\vx'}}{\int_{[0,1)^{d}} f(\vx)^2 d{\vx}-\mu^2} = \frac{I_1}{I_2-(I_3)^2}.
\end{align*}
Therefore, we can define $S_u(\vI)$ and $\overline{S}_u(\vI)$ as functions over a vector of integrals $\vI$. If we can estimate $\vI$ such that $\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI}):=[\widehat{\vI}-\varepsilon_{\vI},\widehat{\vI}+\varepsilon_{\vI} ]$, then we can define the following estimators for the Sobol' indices:
\begin{align*}
\widehat{S}_u & = \frac{1}{2}\left(\min\left(\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),1\right) + \max\left(\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),0\right) \right) \\
\varepsilon_{S_u} & = \frac{1}{2}\left(\min\left(\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),1\right) - \max\left(\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} S_u(\vI),0\right) \right)
\end{align*}
where the true value $S_u\in [\widehat{S}_u - \varepsilon_{S_u}, \widehat{S}_u + \varepsilon_{S_u}]$.

\section{Recursive estimation procedure}
\label{section.recursive.procedure}

The recursive estimation procedure we propose combine the error bound presented in the previous section with one of the two estimation strategies of section \ref{estimation.strategies}. Namely, Saltelli's strategy and the one based on replicated designs. The form of the error bound is the same for both strategies and is given by equation \ref{}.

We start be detailing our recursive procedure in the form of an algorithm. Then, we discuss a possible improvement by considering an estimator recently introduced by Owen improving the estimation of small first-order indices.

\subsection{Main algorithm}
%
%$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each procedure.
%\begin{algorithm}[!ht]
%\caption{Procedure $A$}
%\begin{algorithmic}[1]
%\vspace*{0.2cm}
%\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
%\While {stopping criterion}
%\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$
%
%\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
%\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
%\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=2^{\ell-1}}^{2^{\ell}-1}$
%\State Compute  $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec})
%\EndFor
%\State $\ell \leftarrow \ell + 1$
%\EndWhile
%\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
%\end{algorithmic}
%\label{procedureA}
%\end{algorithm}
%\begin{algorithm}[!ht]
%\caption{Procedure $B$}
%\begin{algorithmic}[1]
%\vspace*{0.2cm}
%\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
%\While {stopping criterion}
%\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$
%
%\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
%\State Compute $\{f(\vx_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\For {$u \in \mathcal{D}$}
%\State  order $\{f({\vx'}_i)\}_{i=2^{\ell-1}}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=2^{\ell-1}}^{2^\ell-1}$
%\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ with formula (\ref{first.order.est.rec})
%\EndFor
%\State $\ell \leftarrow \ell + 1$
%\EndWhile
%\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
%\end{algorithmic}
%\label{procedureB}
%\end{algorithm}
%
\subsection{Improvements}

We focus on the use of a better estimator to evaluate small first-order Sobol' indices in Variant $A$. This estimator is called ``Correlation 2" and was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices. Our aim is to show that using ``Correlation 2" in procedure $A$ may reduce the total number of model evaluations. Its formula writes as follows:
\begin{equation}
\widehat{\underline{\tau}_u^2} = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i)),
\label{correlation2}
\end{equation}
It uses an extra set of $n$ model evaluations to estimate $\widehat{\underline{\tau}_u^2}$. %Owen explains the effectiveness of (\ref{correlation2}) as follows. If the main effect of $\vx_u$ is small then $f(\vx)$ will be almost entirely explained by $\vx_{-u}$. The two terms $(f(\vx_i)-f({\vz}_{i,u}:{\vx}_{i,-u}))$ and $(f(\vx_{i,u}:{\vx'}_{i,-u})-f({\vx'}_i))$ should then be small quantities as well as (\ref{correlation2}). 
\bigskip

We discuss now the potential improvement brought by the use of ``Correlation 2" in Variant $A$. The idea is to replace the current estimator (\ref{first.order.est}) by (\ref{correlation2}) for each small first-order index. Assume that the number of small first-order indices is known and equals $\gamma$. We denote by $u_1,\dots,u_{\gamma}$ the indexes of the corresponding inputs and $\Gamma = \{1,\dots,\gamma\}$. The cost of Variant $A$ including ``Correlation 2" writes: 
\begin{equation}
2 \sum \limits_{j \in \Gamma} n'_{u_j} + \sum \limits_{j \in \mathcal{D}/\Gamma} n'_{u_j} + 2 \ n'_{\max},
\label{cost.improvement}
\end{equation}
where: \begin{itemize}
\item[$\bullet$] $2 \sum \limits_{j \in \Gamma} n'_{u_j}$ is the number of evaluations induces by the $\gamma$ sets \\ $\{f\left(\vz_{i,u_j}:\vx_{i,-u_j}\right)\}_{i=0}^{n'_{u_j}-1}$ and $\gamma$ sets $\{f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)\}_{i=0}^{n'_{u_j}-1}$,
\item[$\bullet$] $\sum \limits_{j \in \mathcal{D}/\Gamma} n'_{u_j}$ is the number of evaluations induced by the $d-\gamma$ sets \\ $\{f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)\}_{i=0}^{n'_{u_j}-1}$, 
\item[$\bullet$] $2 \ n'_{\max}=2 \ \max \limits_{j \in \mathcal{D}} \ (n'_{u_j})$, is the number of evaluations induced by the $2$ sets $\{f\left(\vx_{i}\right)\}_{i=0}^{n'_{\max}-1}$ and $\{f\left({\vx'}_{i}\right)\}_{i=0}^{n'_{\max}-1}$.
\end{itemize}
At the opposite, the cost of Variant $A$ excluding ``Correlation 2" writes: 
\begin{equation}
\sum \limits_{j \in \mathcal{D}} n_{u_j} + 2 \ n_{\max}.
\label{cost.non.improvement}
\end{equation}
The comparison between (\ref{cost.improvement}) and (\ref{cost.non.improvement}) is driven by whether or not total effect indices require more evaluations $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ than their respective first-order indices. We distinguish two extreme cases :
\begin{itemize}
\item[1)] $\forall j \in \Gamma$, the total effect index $\ov{$S$}_j$ requires more evaluations $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ than the first-order index $\underline{S}_j$. As a result, the difference (\ref{cost.improvement}) - (\ref{cost.non.improvement}) is always positive. Thus, using ``Correlation 2" results in a cost regression.
\item[2)] $\forall j \in \Gamma$, the total effect index $\ov{$S$}_j$ requires less evaluations $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ than the first-order index $\underline{S}_j$. As a result, the difference (\ref{cost.improvement}) - (\ref{cost.non.improvement}) is always negative. Thus, using ``Correlation 2" results in a cost improvement.
\end{itemize}
The standard case varies between these two. That is a portion of indices $\ov{$S$}_j$ require more evaluations and another portion of $\ov{$S$}_j$ require less. Illustrations of this potential improvement are addressed in Section \ref{appli}. 
\bigskip

In pratice, one does not know which are the small Sobol' indices. To overcome this issue, we proceed as follows in Variant $A$:
... etc ...

%The second improvement focus on the choice of the $2d$ generating matrices that generate the two Sobol' sequences $\mathcal{P}_\ell$  and ${\mathcal{P}'}_\ell$ in procedure $B$.
%
%We note by $C_1,\dots,C_d$ the matrices generating $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and ${C'}_1,\dots,{C'}_d$ those generating $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$.
%In the multiplicative approach \cite{crass}, these $2d$ matrices are selected as the first $2d$ elements of the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values. 
%
%%In procedure $A$, each design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i\geq0}$ constructed in step $6$ of Algorithm \ref{procedureA} is a $d$-dimensional Sobol' sequence generated by the $d$-subset: $${C'}_1,\dots,{C'}_{u-1},C_u,{C'}_{u+1},\dots,C_d.$$ Therefore, $2$-dimensional projection $t$-values are still optimized.
%
%In procedure $B$, each permutation $\pi_u$ applied in step $6$ of Algorithm \ref{procedureB} results in a design $\{\vx_{i,u}:{\vx'}_{\pi_u(i),-u}\}_{i=0}^{2^\ell-1}$. The $d$ generating matrices corresponding to this design are:
%$${C'}_1U_u,\dots,{C'}_{u-1}U_u,C_u,{C'}_{u+1}U_u,{C'}_dU_u$$ where $U_u$ is a a full rank upper triangular matrix verifying $C_u={C'}_uU_u$. The right multiplication by $U_u$ is the digital representation of the ordering by $\pi_u$.

@@

\subsection{Classical test functions}
